### 2019-03-04 report/final-report.tex
Checkpoint

To evaluate how many user-supplied examples the algorithm requires in
practice, we \textit{randomly} generated appropriate source/target
pairs, mimicking what a na\"{i}ve user might do.  We did not write the
examples by hand out of concern that our knowledge of the synthesis
algorithm might bias the selection. Figure~\ref{fig:exs-reqd} shows
the number of randomly generated examples it takes to synthesize the
correct lens averaged over ten runs.  The synthesis algorithm almost never needs
any examples: only 5 benchmarks need a nonzero number of examples to
synthesize the correct lens and only one, \CF{cust/workitem-probs} required over
10 randomly generated examples.
A clever user may be able to reduce the
number of examples further by selecting examples carefully; we
synthesized \CF{cust/workitem-probs} with only 8 examples.

These numbers are low because there are relatively few well-typed
bijective lenses between any two source and target regular expressions. 
As one would expect, the benchmarks where there are multiple ways to
map source data to the target (and vice versa) require the most examples.
For example, the benchmark \CF{cust/workitem-probs} requires a large number of
examples because it
must differentiate between data in different text fields in both the
source and target and map between them appropriately.  As these text fields are
heavily permuted
(the legacy format ordered fields by a numeric ID, where
the modern format ordered fields alphabetically) and fields can be
omitted, a number of examples are needed to correctly identify the mapping
between fields.

@@

\paragraph*{Comparison Against Other Tools}
%
We are the first tool to synthesize bidirectional transformations between data
formats, so there is no tool to which we can make an apple-to-apples comparison.
Instead, we compare against tools for generating unidirectional
transformations instead. 
Figure~\ref{fig:synthesis-times} includes a comparison against two other
well-known tools that synthesize
text transformation and extraction functions from examples -- Flash Fill and FlashExtract.  For this
evaluation, we used the version of these tools distributed through the
PROSE project~\cite{prose}.

@@

As these tools were designed for a broader audience, they put less of a burden
on the user.  These tools only use input/output examples (for Flash
Fill), or marked text regions (for FlashExtract), as opposed
to \Optician{}'s use of regular expressions to constrain the format of
the input and output.  By using regular expressions,
\Optician{} is able to synthesize significantly more programs
than either existing tool.

Flash Fill and FlashExtract have two tasks: to determine how
the data is transformed, they must also infer the structure of the data, a
difficult job for complex formats.
In particular, neither Flash Fill nor FlashExtract was able to synthesize
transformations or extractions present under two iterations, a type of format
that is notoriously hard to infer.
These types of dual iterations are pervasive in Linux configuration
files, making Flash Fill and FlashExtract ill suited for many of the synthesis
tasks present in our test suite.

Furthermore, as unidirectional transformations, Flash Fill and FlashExtract have
a more expressive calculus.  To guarantee bidirectionality, our syntax must be
highly restrictive, providing a smaller search space to traverse.

@@

We have extended the bijective synthesis tool to synthesize quotient lenses. We will use 
``\QOpt'' to denote our extended version of Optician, and just plain ``Optician'' to denote
the pre-quotient version of Optician. The synthesis algorithm produces Boomerang lens
values, so Boomerang gives synthesized lenses the same first-class status as hand-written ones. 
%
All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
of 1600 MHz DDR3 running macOS High Sierra.

@@

We analyzed the same 39 lens synthesis tasks from the original
Optician system.
We also experimented using our tool to synthesize quotient lenses
between XML, RDF and and JSON formats using data from the data.gov
database; the data consisted of census statistics, demographic
statistics, wage comparosion data, and crime index data).
Recall that to use bijective synthesis,  we had to modify 10 of 39
benchmarks make them bijective (for instance, by normalizing whitespace)
Because quotient lenses are more expressive, we were able avoid such
modifications. This
experience alone points to the benefits that quotient lens synthesis
brings to the table.

\paragraph*{Evaluating programmer effort}

@@

For each problem in the benchmark suite, we calculate the following
measures as proxies for the level of programmer effort when using each
the three approaches:

>>>>

To evaluate how many user-supplied examples the algorithm requires in
practice, we \textit{randomly} generated appropriate source/target
pairs, mimicking what a na\"{i}ve user might do.  (We did not write the
examples by hand out of concern that our knowledge of the synthesis
algorithm might bias the selection.) Figure~\ref{fig:exs-reqd} shows
the number of randomly generated examples it takes to synthesize the
correct lens, averaged over ten runs.  Strikingly, the synthesis algorithm
almost never needs 
any examples at all: only 5 benchmarks need a nonzero number of examples to
synthesize the correct lens and only one required more than
10 randomly generated examples.
(A clever user may be able to reduce the
number of examples further by selecting examples carefully; we were able to
synthesize this last test case with only 8 examples.)

These numbers are low because there are relatively few well-typed
bijective lenses between any two source and target regular expressions. 
As one would expect, the benchmarks where there are multiple ways to
map source data to the target (and vice versa) require the most examples.
For example, one benchmark requires a large number of
examples because it
must differentiate between data in different text fields in both the
source and target and map between them appropriately.  As these text fields are
heavily permuted
(the legacy format ordered fields by a numeric ID, where
the modern format ordered fields alphabetically) and fields can be
omitted, a number of examples are needed to correctly identify the mapping
between fields.

@@

\paragraph*{Comparison Against Other Tools}
%
Ours is the first tool to synthesize bidirectional transformations between data
formats, so there is no tool to which we can make an apple-to-apples comparison.
Instead, we compared against tools for generating unidirectional
transformations. 
Figure~\ref{fig:synthesis-times} includes a comparison against two other
well-known tools that synthesize
text transformation and extraction functions from examples -- Flash Fill and FlashExtract.  For this
evaluation, we used the version of these tools distributed through the
PROSE project~\cite{prose}.

@@

As these tools were designed for a broader audience, they put less of a burden
on the user.  These tools only use input/output examples (for Flash
Fill), or marked text regions (for FlashExtract), as opposed
to our use of regular expressions to constrain the format of
the input and output.  By using regular expressions,
we can synthesize significantly more programs
than either existing tool.

Flash Fill and FlashExtract have two tasks: to determine how
the data is transformed, they must also infer the structure of the data, a
difficult job for complex formats.
In particular, neither Flash Fill nor FlashExtract was able to synthesize
transformations or extractions where the regular expressions describing the
formats involved two nested iteration (Kleene-star) operators---a type of format
that is notoriously hard to infer.
However, this kind of nested iteration is pervasive in Linux configuration
file formats, making Flash Fill and FlashExtract ill suited for many of the
synthesis tasks in our test suite.

Conversely, since they deal with only unidirectional transformations, Flash
Fill and FlashExtract can offer a more expressive set of operators.  To
guarantee bidirectionality, our syntax must be highly restrictive, providing
a smaller search space to traverse.

@@

Our next thread of work was extending the bijective synthesis tool to
synthesize quotient lenses. In this subsection, we will use ``\QOpt'' to
denote our extended version of Optician, and just plain ``Optician'' to
denote the pre-quotient version of Optician. The synthesis algorithm
produces Boomerang lens values, so Boomerang gives synthesized lenses the
same first-class status as hand-written ones.
%
All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
of 1600 MHz DDR3 running macOS High Sierra.

@@

We analyzed the same 39 lens synthesis tasks as we did for the original
Optician system.
We also experimented using our tool to synthesize quotient lenses
between XML, RDF and and JSON formats using data from the data.gov
database; the data consisted of census statistics, demographic
statistics, wage comparosion data, and crime index data).
Recall that, to use bijective synthesis, we had to modify some of the 
benchmarks (10 of 39) to make them bijective---for instance, by normalizing whitespace.
Because quotient lenses are more expressive, we were able avoid such
modifications. This
experience alone points to the benefits that quotient lens synthesis
brings to the table.

\paragraph*{Evaluating Programmer Effort}

@@

For each problem in the benchmark suite, we calculate the following measures
as proxies for the level of programmer effort when using each the three
approaches: \bcp{This seems very detailed, and it includes some technical
  terms (like ``canonize(q)'') that readers of this summary may not
  know.  But maybe it's OK?}
